---
title: "SVI at CBG level"
author: "Derek Ouyang"
date: "10/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(censusapi)
library(tigris)
library(tidyverse)

Sys.setenv(CENSUS_KEY="c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")

acs_vars_2018_5yr <-
  listCensusMetadata(
    name = "2018/acs/acs5",
    type = "variables"
  )
```

```{r}
scc_cbgs <- 
  block_groups("CA","Santa Clara", cb = T, progress_bar = F) %>% 
  select(CBG = GEOID)
```

Official SVI uses a lot of subject tables that don't have CBG level results. I'm replacing everything with the best ACS 5-yr summary version that gives CBG level results. Sometimes there's a slight difference in the "universe" considered. I can document these substitutions more explicitly in the future if needed.

I am not including margins of error. Seems like official SVI includes it just so it's there, but doesn't effect actual scoring.

```{r, eval = F}
scc_cbgs_acs <-
  getCensus(
    name = "acs/acs5",
    vintage = 2018,
    region = "block group:*",
    regionin = "state:06+county:085",
    vars = c(
      "B17010_001E", # Families
      "B17010_002E", # Income in the past 12 months below poverty level
      "B23025_001E", # Population 16 years and over
      "B23025_005E", # Unemployed
      "B19301_001E", # Per capita income in the past 12 months (in 2018 inflation-adjusted dollars)
      "B28006_001E", # Household population 25 years and over
      "B28006_002E", # Less than high school graduate
      "B01001_001E", # Population
      "B01001_020E", # 65 years and over
      "B01001_021E", 
      "B01001_022E", 
      "B01001_023E", 
      "B01001_024E", 
      "B01001_025E", 
      "B01001_044E", 
      "B01001_045E", 
      "B01001_046E", 
      "B01001_047E", 
      "B01001_048E", 
      "B01001_049E", 
      "B01001_003E", # Under 18 years
      "B01001_004E", 
      "B01001_005E", 
      "B01001_006E", 
      "B01001_027E", 
      "B01001_028E", 
      "B01001_029E", 
      "B01001_030E", 
      "C21007_001E", # Civilian population 18 years and over for whom poverty status is determined
      "C21007_005E", # With a disability
      "C21007_008E", 
      "C21007_012E", 
      "C21007_015E",
      "C21007_020E", 
      "C21007_023E",
      "C21007_027E", 
      "C21007_030E",
      "B11003_001E", # Families
      "B11003_010E", # One parent, children under 18 years
      "B11003_016E",
      "B03002_003E", # Not Hispanic or Latino + White alone
      "B16004_001E", # Population 5 years and over
      "B16004_007E", # Speak English less than well
      "B16004_008E", 
      "B16004_012E",
      "B16004_013E",
      "B16004_017E",
      "B16004_018E",
      "B16004_022E",
      "B16004_023E",
      "B16004_029E",
      "B16004_030E",
      "B16004_034E",
      "B16004_035E",
      "B16004_039E",
      "B16004_040E",
      "B16004_044E",
      "B16004_045E",
      "B16004_051E",
      "B16004_052E",
      "B16004_056E",
      "B16004_057E",
      "B16004_061E",
      "B16004_062E",
      "B16004_066E",
      "B16004_067E",
      "B25024_001E", # Housing units
      "B25024_007E", # 10 or more
      "B25024_008E", 
      "B25024_009E", 
      "B25024_010E", # Mobile homes
      "B25014_001E", # Occupied housing units
      "B25014_005E", # Greater than 1 occupant per room
      "B25014_006E", 
      "B25014_007E", 
      "B25014_011E", 
      "B25014_012E", 
      "B25014_013E", 
      "B25044_003E", # No vehicle available
      "B25044_010E",
      "B09019_038E", # In group quarters
      "B27010_001E", # Civilian noninstitutionalized population
      "B27010_017E", # No health insurance coverage
      "B27010_033E", 
      "B27010_050E", 
      "B27010_066E"
    )
  )

saveRDS(scc_cbgs_acs, "scc_cbgs_acs.rds")
```

CBGs with no population
060855045042
060855002002 (Jail)

```{r, eval = F}
scc_svi <-
  scc_cbgs_acs %>% 
  transmute(
    CBG = paste0(state,county,tract,block_group),
    EP_POV = B17010_002E/B17010_001E*100,
    EP_UNEMP = B23025_005E/B23025_001E*100,
    EP_PCI = B19301_001E,
    EP_NOHSDP = B28006_002E/B28006_001E*100,
    EP_AGE65 = (
      B01001_020E +
      B01001_021E +
      B01001_022E +
      B01001_023E +
      B01001_024E +
      B01001_025E +
      B01001_044E +
      B01001_045E +
      B01001_046E +
      B01001_047E +
      B01001_048E +
      B01001_049E
    )/B01001_001E*100,
    EP_AGE17 = (
      B01001_003E +
      B01001_004E +
      B01001_005E +
      B01001_006E +
      B01001_027E +
      B01001_028E +
      B01001_029E +
      B01001_030E
    )/B01001_001E*100,
    EP_DISABL = (
      C21007_005E +
      C21007_008E +
      C21007_012E +
      C21007_015E +
      C21007_020E +
      C21007_023E +
      C21007_027E +
      C21007_030E
    )/C21007_001E*100,
    EP_SNGPNT = (
      B11003_010E +
      B11003_016E
    )/B11003_001E*100,
    EP_MINRTY = (1-B03002_003E/B01001_001E)*100,
    EP_LIMENG = (
      B16004_007E +
      B16004_008E +
      B16004_012E +
      B16004_013E +
      B16004_017E +
      B16004_018E +
      B16004_022E +
      B16004_023E +
      B16004_029E +
      B16004_030E +
      B16004_034E +
      B16004_035E +
      B16004_039E +
      B16004_040E +
      B16004_044E +
      B16004_045E +
      B16004_051E +
      B16004_052E +
      B16004_056E +
      B16004_057E +
      B16004_061E +
      B16004_062E +
      B16004_066E +
      B16004_067E
    )/B16004_001E*100,
    EP_MUNIT = (
      B25024_007E +
      B25024_008E +
      B25024_009E
    )/B25024_001E*100,
    EP_MOBILE = B25024_010E/B25024_001E*100,
    EP_CROWD = (
      B25014_005E +
      B25014_006E +
      B25014_007E +
      B25014_011E +
      B25014_012E +
      B25014_013E
    )/B25014_001E*100,
    EP_NOVEH = (
      B25044_003E +
      B25044_010E
    )/B25014_001E*100,
    EP_GROUPQ = B09019_038E/B01001_001E,
    EP_UNINSUR = (
      B27010_017E +
      B27010_033E +
      B27010_050E +
      B27010_066E
    )/B27010_001E*100
  ) %>% 
  filter(!is.na(EP_POV)) %>% 
  mutate(
    EPL_POV = percent_rank(EP_POV),
    EPL_UNEMP = percent_rank(EP_UNEMP),
    EPL_PCI = percent_rank(EP_PCI),
    EPL_NOHSDP = percent_rank(EP_NOHSDP),
    EPL_AGE65 = percent_rank(EP_AGE65),
    EPL_AGE17 = percent_rank(EP_AGE17),
    EPL_DISABL = percent_rank(EP_DISABL),
    EPL_SNGPNT = percent_rank(EP_SNGPNT),
    EPL_MINRTY = percent_rank(EP_MINRTY),
    EPL_LIMENG = percent_rank(EP_LIMENG),
    EPL_MUNIT = percent_rank(EP_MUNIT),
    EPL_MOBILE = percent_rank(EP_MOBILE),
    EPL_CROWD = percent_rank(EP_CROWD),
    EPL_NOVEH = percent_rank(EP_NOVEH),
    EPL_GROUPQ = percent_rank(EP_GROUPQ),
    EPL_UNINSUR = percent_rank(EP_UNINSUR)
  ) %>% 
  mutate(
    EPL_SUM = 
      EPL_POV +
      EPL_UNEMP +
      EPL_PCI +
      EPL_NOHSDP +
      EPL_AGE65 +
      EPL_AGE17 +
      EPL_DISABL +
      EPL_SNGPNT +
      EPL_MINRTY +
      EPL_LIMENG +
      EPL_MUNIT +
      EPL_MOBILE +
      EPL_CROWD +
      EPL_NOVEH +
      EPL_GROUPQ +
      EPL_UNINSUR,
    SVI = percent_rank(EPL_SUM)*100
  ) %>% 
  select(
    CBG,
    SVI,
    everything(),
    -EPL_SUM
  ) %>% 
  left_join(scc_cbgs) %>% 
  st_as_sf() %>%
  st_transform(4326)

saveRDS(scc_svi, "scc_svi.rds")
write_csv(scc_svi %>% st_set_geometry(NULL), "scc_svi.csv")
```